---
title: "SQL in R"
author: "Linh Tran"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
library(sqldf)
library(RSQLite)
library(RODBC)


```


```{r}
sqldf('SELECT age, circumference 
      FROM Orange 
      WHERE Tree = 1 
      ORDER BY circumference ASC')
```

# SQL Queries

SQL syntax is case insensitive so the two following commands are equivalent

```{r}
sqldf("SELECT * FROM iris")
sqldf("select*from iris")


# Basic syntax: SELECT variable 1, variable 2 FROM DATA
```


```{r}
data(BOD)
head(BOD)

sqldf('SELECT demand FROM BOD')
sqldf('select time,demand from BOD')

```


If you have variable with `.` in their name, refer to them in quotes

```{r}
sqldf('SELECT "Petal.Width" FROM iris')
```

## Wildcard

A wild card can be passed to extract everything

```{r}
bod2 <- sqldf('SELECT * FROM BOD')
bod2[1:3,]
```


## LIMIT

To control the number of results returned, use `LIMIT #`

```{r}
sqldf('SELECT * FROM iris LIMIT 5')
```


## ORDER BY

To order variables, use the syntax `ORDER BY var1 {ASC/DESC}, var2 {ASC/DESC}` where the choice of `ASC` for ascending or `DESC` for descending is made per variable

```{r}
sqldf ("SELECT * FROM Orange ORDER BY age ASC, circumference DESC LIMIT 10")
```


## WHERE

Conditional statemenets can be added via `WHERE`

```{r}
sqldf ("SELECT demand from BOD WHERE Time <6")
```

Both `AND` and `OR` are valid, along with parentheses to affect order of operations

```{r}
sqldf("SELECT * FROM rock WHERE (peri > 5000 AND shape < 0.05) OR perm > 1000")
```


### `IN`

`WHERE IN ` is used similar to R's ` %in% `. It also supports `NOT`. 

```{r}
sqldf("SELECT * FROM BOD WHERE Time IN (1,7)")

sqldf("SELECT * FROM BOD WHERE Time NOT IN (1,7)")

sqldf("SELECT * FROM BOD WHERE (Time > 1 AND Time < 7)")

```


### `LIKE`

`LIKE` can be thought of as a weak regular expression command. It only allows the single wildcard `%` which matches any number of charactesr. For example, to extract the data where the feed ends with "bean":

```{r}
sqldf('SELECT * FROM chickwts WHERE feed LIKE "%bean" LIMIT 5')
```



## Aggregated data

Select statements can create aggregated data using `AVG`, `MEDIAN`, `MAX`, `MIN`, `SUM` as functions in the list of variables to select. The `GROUP BY` statement can be added to aggregate by groups. `AS` to rename

```{r}
sqldf('SELECT AVG(circumference) FROM Orange')
```


```{r}
sqldf('SELECT tree, AVG(circumference) AS meancirc FROM Orange GROUP BY tree')

```


### Counting data

`SELECT COUNT()` retunrs the number of observations. Passing `*` or nothing returns total rows, passing a variable name returns the number of non-NA entries. `AS` works as well.

```{r}
d <- data.frame(a = c(1,1,1), b = c(1, NA, NA))

d
```

```{r}
sqldf('SELECT COUNT() as numrows FROM d')

sqldf('SELECT COUNT(a) from d')
```

